---
title: "anc-pregnancy-loss-lbw"
author: "Allan Kimaina"
output:
  Grmd::docx_document:
    fig_caption: TRUE
    force_captions: TRUE
---

<style type="text/css">
body {
  max-width: 100em !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      echo=FALSE,
                      dpi=96,
                      fig.width=4, fig.height=4, # Default figure widths
                      dev="png", #dev.args=list(type="cairo"), # The png device
                      # Change to dev="postscript" if you want the EPS-files
                      # for submitting. Also remove the dev.args() as the postscript
                      # doesn't accept the type="cairo" argument.
                      error=FALSE)
 
# Evaluate the figure caption after the plot
knitr::opts_knit$set(eval.after='fig.cap')
 
# Use the table counter that the htmlTable() provides
options(table_counter = TRUE)
 
# Use the figCapNo() with roman letters
options(fig_caption_no_roman = TRUE)
#options(kableExtra.latex.load_packages = F)

# Then install the Grmd-package by running below code:
#devtools::install_github("gforge/Grmd")

# function to install missing packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, repos='http://cran.rstudio.com/')
  sapply(pkg, require, character.only = TRUE)
}

#install.packages('package_name', dependencies=TRUE, repos='http://cran.rstudio.com/')

packages =c( "dplyr",  "readxl","Hmisc","Gmisc", "magrittr", "flextable", "MASS", "tidyverse")

ipak(packages)


select = dplyr::select; summarize = dplyr::summarize; rename = dplyr::rename; mutate = dplyr::mutate;

#names(dataset)
```


## Utils
```{r warning=FALSE}

describeMissing= function (x, html = TRUE, number_first = TRUE, percentage_sign = TRUE, 
    language = "en", useNA.digits = 1, ...) {
    if (!any(is.na(x))) 
        return(invisible())
    df_arg_list <- list(x = is.na(x), html = html, number_first = number_first, 
        percentage_sign = percentage_sign, language = language, 
        digits = useNA.digits)
    dot_args <- list(...)
    for (n in names(dot_args)) {
        if (!n %in% names(df_arg_list)) {
            df_arg_list[[n]] <- dot_args[[n]]
        }
    }
    missing <- fastDoCall(describeFactors, df_arg_list)
   rownames(missing)<-c("FALSE","Missing")
    return(missing["Missing", ])
}

desc_both <- function(x, ...) {
    result <- c(
      describeMean(x, useNA="no"),
      describeMedian(x, useNA="no"),
      describeMissing(x)
    )
    return(result)
}

desc_mean <- function(x, ...) {
    result <- c(
      describeMean(x, useNA="no"),
      describeMissing(x)
    )
    return(result)
 }
 

desc_median <- function(x, ...) {
    result <- c(
      describeMedian(x, useNA="no"),
      describeMissing(x)
    )
    return(result)
 }

 
 
# Creating a wrapper for getting descriptive statistics
getTable1Stats <- function(x, y, cont_fx=desc_both, data=dataset, digits = 1,statistics = T){
  data=data%>%drop_na(all_of(y))
  getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = T,
                        total_col_show_perc = T,
                        header_count = TRUE)
  
}

relevelBy <- function(varList, data){
  for (factorName in names(varList)) {
    for(var in varList[[factorName]]){
      data[[var]] =relevel(factor(data[[var]]),factorName)
    }
  }
  return(data)
}
```

## READ and CLEAN

```{r warning=F}

dataset <- readxl::read_excel("merged.xlsx")%>%
              mutate(low_birth_weight=factor(ifelse(birth_weight_grams<2500,"Yes", "No")),
                     
                     # outcome=ifelse(preg_outcome=="LTFU", "LTFU", NA),
                     # outcome=factor(ifelse(is.na(outcome), "Completed",outcome)),
                    
                     preg_loss=ifelse(preg_outcome=="Twins"|preg_outcome=="Preterm", "No",preg_outcome),
                     preg_loss=ifelse(is.na(preg_loss), "No",preg_loss),
                     preg_loss=ifelse(preg_loss=="LTFU", NA, preg_loss),
                     preg_loss=factor(ifelse(preg_loss=="No", "No", "Yes")),
                     
                    
                     preterm=ifelse(pregnancy_duration_weeks<=37 & preg_loss=="No", "Yes","No"),
                     
                    
                     outcome=ifelse(preg_loss=="Yes", "Pregnancy Loss", "Delivered"),
                     outcome=factor(ifelse(is.na(outcome), "LTFU",outcome)),
                     
                     
                   
                     
                     # SE
                     years_marriage=ifelse(cp_marital_status=="Married",sh_relationship_years,NA),
                     years_single=ifelse(cp_marital_status=="Single",sh_relationship_years,NA),
                     polygamous=ifelse(sh_relationship_type=="Polygamous","Yes","No"),
                     education_level=ifelse(se_education_level=="College"|se_education_level=="University","College/University",se_education_level),
                     
                     # ANC
                     trimester_1st_anc=ifelse(cp_anc_1st_anc<=12,"First (<= 12 weeks)",NA),
                     trimester_1st_anc=ifelse( cp_anc_1st_anc>12&cp_anc_1st_anc<=24,"Second (13 – 24)",trimester_1st_anc),
                     trimester_1st_anc=ifelse( cp_anc_1st_anc>24,"Third (>24 weeks)",trimester_1st_anc),
                     
                     anc_hb_low=ifelse(gesthb<=12&ancp_hb_results<11.0,"Yes", NA),
                     anc_hb_low=ifelse(gesthb>24&ancp_hb_results<11.0,"Yes", anc_hb_low),
                     anc_hb_low=ifelse(gesthb>12&gesthb<=24&ancp_hb_results<10.5,"Yes", anc_hb_low),
                     anc_hb_low=relevel(factor(ifelse(is.na(anc_hb_low)&!is.na(ancp_hb_results), "No",anc_hb_low)),"Yes"),
                     
                     
                     hiv_haart_regimen=relevel(factor(ifelse(hiv_haart_regimen=="Yes","Yes","N/A")),"Yes"),
                     ancp_vdrl_rpr_results=relevel(factor(ifelse(ancp_vdrl_rpr_results=="NR","Negative",ancp_vdrl_rpr_results)),"Positive"),
                     ancp_hiv_results=relevel(factor(ifelse(ancp_hiv_results=="NR","Negative","Positive")),"Positive"),
                     
                     
                     # Lab
                     ferritin_cat=ifelse(ferritin<=15,"<= 15",NA),
                     ferritin_cat=ifelse( ferritin>16&ferritin<=30,"16 – 30",ferritin_cat),
                     ferritin_cat=ifelse( ferritin>30,">30",ferritin_cat),
                     
                     lab_hb_low=ifelse(gesthb<=12&lab_hgb<11.0,"Yes", NA),
                     lab_hb_low=ifelse(gesthb>24&lab_hgb<11.0,"Yes", lab_hb_low),
                     lab_hb_low=ifelse(gesthb>12&gesthb<=24&lab_hgb<10.5,"Yes", lab_hb_low),
                     lab_hb_low=relevel(factor(ifelse(is.na(lab_hb_low)&!is.na(lab_hgb), "No",lab_hb_low)),"Yes"),
                     
                     
                     
                       # FIX 
                     #inpouch=relevel( factor(ifelse(is.na(inpouch), "Missing",inpouch)), "Negative" ),
                     cp_sbp=as.numeric(cp_sbp),
                     inpouch=ifelse(is.na(inpouch), "Missing",inpouch),
                     years_single=ifelse(code_no=="LB325",-0.1,years_single),
  
                     
                     preg_outcomexx=preg_outcome
              )
                     
varList= list(
"Yes"=c('polygamous','anc_hb_low','hiv_haart_regimen','lab_hb_low'),
"Positive"=c('ancp_rhesus_results','ancp_vdrl_rpr_results','ancp_hiv_results','inpouch','lab_rdt','chlamydia','gonorrhea')
)



dataset= relevelBy(varList,dataset) 
  
```


# Univariate Analysis




## Table 1: Demographics and socio-economic characteristics
```{r warning=FALSE}

# Getting descriptive statistics 
mergeDesc("Age (years)" = getTable1Stats("age","outcome",desc_median,statistics=T),
          "Marital Status" = getTable1Stats("cp_marital_status","outcome",desc_both,statistics=T),
          "Years of marriage" =getTable1Stats("years_marriage","outcome",desc_median,statistics=T),  # TODO delete missing from excel
          "Years of relationship (single)" = getTable1Stats("years_single","outcome",desc_median,statistics=T), # TODO delete missing from excel
          "Polygamous marriage" =getTable1Stats("polygamous","outcome",desc_median,statistics=T),
          "Occupation (%)" = getTable1Stats("se_occupation","outcome",desc_both,statistics=T),
          "Education (%)" =getTable1Stats("education_level","outcome",desc_median,statistics=T)
          )%>%
   htmlTable( caption  = "<b> Demographics and socio-economic characteristics </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1, 3,1 ), cgroup = c('', 'General Study Outcomes', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```


## Table 2: Clinical, ANC profile and previous pregnancy characteristics

```{r warning=FALSE}



# Getting descriptive statistics 
mergeDesc("Weight (Kg)" = getTable1Stats("cp_weight","outcome",desc_median,statistics=T),
          "Height (cm)" = getTable1Stats("cp_height","outcome",desc_median,statistics=T),
          "BMI (Kg/m2)" =getTable1Stats("cp_bmi","outcome",desc_median,statistics=T),
          "Systolic BP (mm Hg)" = getTable1Stats("cp_sbp","outcome",desc_median,statistics=T),
          "Diastolic BP (mm Hg) " =getTable1Stats("cp_dbp","outcome",desc_median,statistics=T),
          "Gestation at enrollment (Weeks)" = getTable1Stats("cp_gbd","outcome",desc_median,statistics=T),
          "Gravidity" =getTable1Stats("gravida","outcome",desc_median,statistics=T),
          "#TODO#Previous abortions [n, (%)]" = getTable1Stats("chlamydia","outcome",desc_both,statistics=T), # TODO
          "Number of ANC visits (%)" = getTable1Stats("cp_anc_attendance","outcome",desc_median,statistics=T),
          "Gestation at first ANC attendance (Weeks)" = getTable1Stats("cp_anc_1st_anc","outcome",desc_median,statistics=T),
          "Trimester at first ANC attendance (%)" = getTable1Stats("trimester_1st_anc","outcome",desc_median,statistics=T),
          "Hemoglobin (g/dL) (ANC)" = getTable1Stats("ancp_hb_results","outcome",desc_median,statistics=T),
          "HB <11.0 at 1st & 3rd trimesters & <10.5 at 2nd trimester" = getTable1Stats("anc_hb_low","outcome",desc_median,statistics=T),
          "GBD at HB measurement (Weeks)" =getTable1Stats("gesthb","outcome",desc_median,statistics=T),
          "Blood Group (%)" = getTable1Stats("ancp_bloodgroup_results","outcome",desc_median,statistics=T),
          "Rhesus" = getTable1Stats("ancp_rhesus_results","outcome",desc_median,statistics=T),
          "VDRL" = getTable1Stats("ancp_vdrl_rpr_results","outcome",desc_median,statistics=T),
          "HIV" = getTable1Stats("ancp_hiv_results","outcome",desc_median,statistics=T),
          "On HAART " = getTable1Stats("hiv_haart_regimen","outcome",desc_median,statistics=T),
          "#TODO#Urinalysis" = getTable1Stats("ancp_urinalysis_results","outcome",desc_median,statistics=T) # TODO
          )%>%
   htmlTable( caption  = "<b> Clinical, ANC profile and previous pregnancy characteristics</b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1, 3,1 ), cgroup = c('', 'General Study Outcomes', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 




```


## Table 3: Laboratory results
```{r warning=FALSE}


# Getting descriptive statistics 
mergeDesc("Ferritin (µg/L)" = getTable1Stats("ferritin","outcome",desc_median,statistics=T),
          "Ferritin (µg/L) (%)" = getTable1Stats("ferritin_cat","outcome",desc_median,statistics=T),
          "Hemoglobin (g/dL) (Lab)" = getTable1Stats("lab_hgb","outcome",desc_median,statistics=T),  
          "HB <11.0 at 1st & 3rd trimesters & <10.5 at 2nd trimester (%)" =getTable1Stats("lab_hb_low","outcome",desc_median,statistics=T),
          "HCT" = getTable1Stats("lab_hct","outcome",desc_median,statistics=T),
          "White Blood cell count" =getTable1Stats("lab_wbc_count","outcome",desc_median,statistics=T),
          "MCV " = getTable1Stats("lab_mcv","outcome",desc_median,statistics=T),
          "Sickle cell carriers" =getTable1Stats("sickle_cell","outcome",desc_median,statistics=T), # TODO
          "Trichonomiasis " =getTable1Stats("inpouch","outcome",desc_median,statistics=T),
          "Vaginosis" =getTable1Stats("vaginosis","outcome",desc_median,statistics=T),
          "Malaria " =getTable1Stats("lab_rdt","outcome",desc_median,statistics=T),
          "Chlamydia " =getTable1Stats("chlamydia","outcome",desc_median,statistics=T),
          "Gonorrhea " =getTable1Stats("gonorrhea","outcome",desc_median,statistics=T)
          )%>%
   htmlTable( caption  = "<b>  Laboratory results </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1, 3,1 ), cgroup = c('', 'General Study Outcomes', ''),tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```



## Table 4: Pregnancy outcomes 
```{r warning=FALSE}

# Getting descriptive statistics 
mergeDesc("Pregnancy duration (weeks)" = getTable1Stats("ferritin","outcome",desc_both,statistics=T),
          "Pregnancy duration <38 weeks (%)" = getTable1Stats("ferritin","outcome",desc_both,statistics=T), 
          "Still Birth or death" = getTable1Stats("lab_hgb","outcome",desc_both,statistics=T),  
          "Delivery mode (%)" =getTable1Stats("lab_hgb","outcome",desc_median,statistics=T), 
          "Birth weight (grams)" = getTable1Stats("lab_hct","outcome",desc_both,statistics=T),
          "Birth weight < 2500 grams" =getTable1Stats("lab_wbc_count","outcome",desc_median,statistics=T),
          "Delivery place (%) " = getTable1Stats("lab_mcv","outcome",desc_both,statistics=T),
          "Delivery by (%)" =getTable1Stats("sickle_cell","outcome",desc_median,statistics=T)
          )%>%
   htmlTable( caption  = "<b> Pregnancy outcomes </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1, 3,1 ), cgroup = c('', 'General Study Outcomes', ''), tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```


## Table 5: Associations with pregnancy loss 
```{r warning=FALSE}

mergeDesc("Maternal age (%)" = getTable1Stats("age","preg_loss",desc_both,statistics=T),
          "BMI" = getTable1Stats("ferritin","preg_loss",desc_both,statistics=T), 
          "Previous abortion " = getTable1Stats("lab_hgb","preg_loss",desc_both,statistics=T),
          "Gravida (%)" =getTable1Stats("lab_hgb","preg_loss",desc_median,statistics=T),
          "Iron deficiency ( Ferritin <= 15 )" = getTable1Stats("lab_hct","preg_loss",desc_both,statistics=T),
          "Anemic " = getTable1Stats("lab_mcv","preg_loss",desc_both,statistics=T),
          "Sickle cell disease" =getTable1Stats("sickle_cell","preg_loss",desc_median,statistics=T),
          "Trichonomiasis" =getTable1Stats("inpouch","preg_loss",desc_median,statistics=T),
          "Vaginosis" =getTable1Stats("vaginosis","preg_loss",desc_median,statistics=T),
          "Chlamydia " =getTable1Stats("chlamydia","preg_loss",desc_median,statistics=T),
          "Gonorrhea " =getTable1Stats("gonorrhea","preg_loss",desc_median,statistics=T),
          "Malaria " =getTable1Stats("lab_rdt","preg_loss",desc_median,statistics=T),

          "HIV " =getTable1Stats("chlamydia","preg_loss",desc_median,statistics=T),
          "Any acute infection " =getTable1Stats("gonorrhea","preg_loss",desc_median,statistics=T),
          "Urinary tract infection  " =getTable1Stats("lab_rdt","preg_loss",desc_median,statistics=T),

          "Syphilis screening" =getTable1Stats("gonorrhea","preg_loss",desc_median,statistics=T),
          "Anemia" =getTable1Stats("lab_mcv","preg_loss",desc_median,statistics=T)
          )%>%
   htmlTable( caption  = "<b>  Associations with pregnancy loss </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1, 2,1 ), cgroup = c('', 'Pregnancy Loss', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```



## Table 6: Associations with low birth weight 
```{r warning=FALSE}

# Getting descriptive statistics 
mergeDesc("Maternal age (%)" = getTable1Stats("age","low_birth_weight",desc_both,statistics=T),
          "BMI" = getTable1Stats("ferritin","low_birth_weight",desc_both,statistics=T), 
          "Previous abortion " = getTable1Stats("lab_hgb","low_birth_weight",desc_both,statistics=T),  
          "Gravida (%)" =getTable1Stats("lab_hgb","low_birth_weight",desc_median,statistics=T), 
          "Iron deficiency ( Ferritin <= 15 )" = getTable1Stats("lab_hct","low_birth_weight",desc_both,statistics=T),
          "Anemic " = getTable1Stats("lab_mcv","low_birth_weight",desc_both,statistics=T),
          "Sickle cell disease" =getTable1Stats("sickle_cell","low_birth_weight",desc_median,statistics=T),
          "Trichonomiasis " =getTable1Stats("inpouch","low_birth_weight",desc_median,statistics=T),
          "Vaginosis" =getTable1Stats("vaginosis","low_birth_weight",desc_median,statistics=T),
          "Chlamydia " =getTable1Stats("chlamydia","low_birth_weight",desc_median,statistics=T),
          "Gonorrhea " =getTable1Stats("gonorrhea","low_birth_weight",desc_median,statistics=T),
          "Malaria " =getTable1Stats("lab_rdt","low_birth_weight",desc_median,statistics=T),
          
          "HIV " =getTable1Stats("chlamydia","low_birth_weight",desc_median,statistics=T),
          "Any acute infection " =getTable1Stats("gonorrhea","low_birth_weight",desc_median,statistics=T),
          "Urinary tract infection  " =getTable1Stats("lab_rdt","low_birth_weight",desc_median,statistics=T),
          
          "Syphilis screening" =getTable1Stats("gonorrhea","low_birth_weight",desc_median,statistics=T),
          "Anemia" =getTable1Stats("lab_mcv","low_birth_weight",desc_median,statistics=T)
          )%>%
   htmlTable( caption  = "<b> Associations with low birth weight   </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1, 2,1 ), cgroup = c('', 'Low Birth Weight', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```


## Table 7: Correlates of birth outcomes
```{r warning=FALSE}

getMultipleTable1Stats <- function(x, y,y2,y3, cont_fx=desc_both, data=dataset, digits = 1,statistics = T){

 yy1= getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = T,
                        total_col_show_perc = T,
                        header_count = TRUE)
 
 yy2= getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y2]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = T,
                        total_col_show_perc = T,
                        header_count = TRUE)
 
 yy3= getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y3]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = T,
                        total_col_show_perc = T,
                        header_count = TRUE)
  
  cbind(yy1,yy2,yy3)
  
}

# Getting descriptive statistics 

mergeDesc(
          "Marital status" = getMultipleTable1Stats("cp_marital_status","low_birth_weight","preterm","preg_loss",desc_both,statistics=T),
          "BMI" = getMultipleTable1Stats("ferritin","low_birth_weight","preterm","preg_loss",desc_both,statistics=T), 
          "Booking trimester" = getMultipleTable1Stats("lab_hgb","preterm","low_birth_weight","preg_loss",desc_both,statistics=T),  
          "Education level" =getMultipleTable1Stats("lab_hgb","low_birth_weight","preterm","preg_loss",desc_median,statistics=T), 
          "Occupation" = getMultipleTable1Stats("lab_hct","low_birth_weight","preterm","preg_loss",desc_both,statistics=T),
          "IPT" = getMultipleTable1Stats("lab_mcv","low_birth_weight","preterm","preg_loss",desc_both,statistics=T),
          "Iron Supplementation" =getMultipleTable1Stats("sickle_cell","low_birth_weight","preterm","preg_loss",desc_median,statistics=T),
          "Deworming" =getMultipleTable1Stats("inpouch","low_birth_weight","preterm","preg_loss",desc_median,statistics=T),
          "Contraceptive use" =getMultipleTable1Stats("vaginosis","low_birth_weight","preterm","preg_loss",desc_median,statistics=T)
          ) %>%
   htmlTable( caption  = "<b> Correlates of birth outcomes  </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,2,1,1,2,1,1,2,1 ), cgroup = c( '','Low Birth Weight', '','','Preterm', '','','Pregnancy Loss', '')
               ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```

# Multivariate Analysis

https://rpubs.com/CaperKen/198209
https://strengejacke.wordpress.com/2014/02/20/no-need-for-spss-beautiful-output-in-r-rstats/

```{r warning=FALSE}
#naniar::gg_miss_var(dataset )+ labs(title = "Frequency of Missingness")

```


```{r warning=FALSE}
variables <- list(
  "ferritin"="ferritin",
  "sickle_cell"="sickle_cell",
  "lab_hgb"="lab_hgb",
  "lab_hct"="lab_hct",                        
  "lab_wbc_count"="lab_wbc_count",                 
  "lab_mcv"="lab_mcv",                       
  "inpouch"="inpouch",                        
  "vaginosis"="vaginosis",                      
  "lab_rdt"="lab_rdt",                        
  "chlamydia"="chlamydia",                     
  "gonorrhea"="gonorrhea"
)

y <- "low_birth_weight"
x <- names(variables)

# fully parameterized!
f <- as.formula(paste(y, paste(x, collapse = " + "), sep = " ~ "))
print(f)

no.na.data <- na.omit(dataset[c(y, x)])

# full model
full.logit = glm(f,data=no.na.data,family=binomial)
summary(full.logit)



# step AIC
MASS::stepAIC(full.logit,direction="both")


```

\newpage
